### 2022-02-18 misc/solutions/prob-method-sols.tex
2.7

\def\given{\,|\,}
\def\biggiven{\,\big|\,}

\maketitle{The Probabilistic Method}{exercise solutions by}{Marcel K. Goh}{15 February 2022}

@@

>>>>

\def\given{\,|\,}
\def\biggiven{\,\big|\,}
\def\F{{\cal F}}

\maketitle{The Probabilistic Method}{exercise solutions by}{Marcel K. Goh}{17 February 2022}

@@

\proclaim Exercise 2.7. Let $\F$ be a family of subsets of $[n] = \{1,2,\ldots,n\}$ and suppose there are
no $A,B\in \F$ satisfying $A\subseteq B$. Let $\sigma\in S_n$ be a random permutation of the elements of
$[n]$ and consider the random variable
$$ X = \bigl| \bigl\{ i : \{\sigma(1),\sigma(2),\ldots,\sigma(i)\}\in \F\bigr\}\bigr|.$$
By considering the expectation of $X$, prove that $|\F|\le {n\choose \lfloor n/2\rfloor}$.

\proof This is the proof that I said Exercise~1.8 reminded me of. Let $S_i = \{\sigma(1),\ldots,\sigma(i)\}$
and let
$$ C = \{\emptyset, S_1, S_2,\ldots, S_n\}.$$
If there are $i<j$ such that $S_i$ and $S_j$ are both in $\F$, then
$\F$ violates the condition that there are no
$A,B\in \F$ with $A\subseteq B$. So $X\le 1$.
On the other hand, $C$ contains
exactly one set of size $i$ for each $0\le i\le n$ and the probability that $S_i$ is a given
element of $X^{(i)}$ is the same. So letting $N_i$ be the number of sets of size $i$ in $\F$, we have
$$1\ge \ex\{X\} = \sum_{A\in \F} {n\choose |A|}^{-1} = \sum_{i=0}^n N_i {n\choose i}^{-1}\ge
{|\F|\over {n\choose \lfloor n/2\rfloor}},$$
where the last inequality holds because ${n\choose k}$ is maximised when $k=\lfloor n/2\rfloor$
and because the sum of the $N_i$ is $|\F|$.\slug
